---
title: "Making Modules"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making Modules}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction 

This vignette will provide an overview of the formods framework for creating reproducable modules that interact with each other. Each module has its own namespace that is mantained by using a module short name as a prefix for functions. For example the figure generation module uses `FG`. If you want to create a module, please submit an issue at the
[formods github repository](https://github.com/john-harrold/formods/issues) with the following information:

* The short name you intend to use 
* A brief description of the module
* A list of modules it depends on

## Current modules

* ASM - [App State Management](https://formods.ubiquity.tools)
* DW - [Data Wrangling](https://formods.ubiquity.tools)
* FG - [Figure Generation](https://formods.ubiquity.tools)
* FM - [Formods](https://formods.ubiquity.tools)
* UD - [Upload Data](https://formods.ubiquity.tools)

The current modules are in development:

* NCA - [Non-Compartmental Analysis](https://github.com/john-harrold/ruminate)

# formods framework

To get started you can make a copy of some example files:

``` r
copy.file(system.file(package="formods", "templates", "ZZ_module_components.R"), "ZZ_module_components.R")
copy.file(system.file(package="formods", "templates", "ZZ_Server.R"), "ZZ_Server.R")
copy.file(system.file(package="formods", "templates", "ZZ.yaml"), "ZZ.yaml")
``` 

- `ZZ_module_components.R` - An app that can be used for testing the module and highlighting the different UI elements that are used within the module. 
- `ZZ_Server.R`: A bare bones file containing the expected functions and their minimal inputs.
- `ZZ.yaml`: This module configuration file contains the the minimal elements expected, but you can add your own fields to suit your modules needs. 

You can rename the `ZZ` portion of the files with the module short name you're using. The `ZZ_Server.R` would go in the `R` directory in your package and the other two files would go into `inst/templates`.

state fields

- `yaml`
- `MC`



# Expected functions


- `ZZ_Server`  
- `ZZ_init_state`  
- `ZZ_fetch_state`  
- `ZZ_fetch_code` This takes as it's first argument the module state. When called with only this argument it should return a character object containing all of the code needed to generate the elements of this module represented in the app. You can assume that any modules this one depends on will be defined previously. For example the `FG` module will return only that code associated with generating figures. It will be appended to the code from the `UD` and `DW` modules that define loading the dataset and creation of the data views. For modules where no code is generated (e.g. `ASM`) just return `NULL`.

# Expected UI elements

- `ui_zz_compact` 

# Helper functions in formods

- `FM_le()`
- `FM_tc()`
- `has_changed()`
- `set_hold()`
- `fetch_hold()`
- `remove_hold()`
- `FM_init_state()`
- `FM_set_mod_state()`
- `FM_fetch_mod_state()`
- `FM_set_ui_msg()`

# Module interaction

Say you are using the UD module to feed data into the DW module and the user goes back to the upload form and uploads a different data set. This will need to trigger a reset of the Data Wrangling module as well as tell your larger app that something has changed. 

## Module state and reacting to changes

 `react_state[[id_UD]][["DS"]][["checksum"]]`

* `isgood` A boolean variable to indicate the general state of the module. 
* `checksum` A checksum that can be used to detect changes in this module. For
  example in the UD module this will change if the uploaded file or the sheet
  selected from a currently uploaded file changes. 
* `code_previous` Carries over the code from other modules the current one
  depends on. For ecample the DW module would depend on the UD model to load
  the data. So `code_previous` here would contain the code to load the
  dataset.
* `code` Code to repelicate the current state of the module. 

## Configuration file
```{css, echo=FALSE}
.scroll-100 {
  max-height: 100px;
  overflow-y: auto;
  background-color: inherit;
}
```



# YAML configuration files {.tabset}

```{r echo=FALSE, message=FALSE, warning=FALSE, eval=TRUE, class.output="scroll-100", comment=''}
#yaml= file.path(system.file(package="formods"), "templates", "formods.yaml")
#cat(readLines(file.path(system.file(package="formods"), "templates", "formods.yaml")), sep="\n")

#library(shiny)
#library(shinyAce)
# renderUI({
# aceEditor("formods", value=readLines(file.path(system.file(package="formods"), "templates", "formods.yaml")))
# })

# yamls = list(
#   formods.yaml= file.path(system.file(package="formods"), "templates", "formods.yaml"),
#   ASM.yaml    = file.path(system.file(package="formods"), "templates", "ASM.yaml"),
#   DS.yaml     = file.path(system.file(package="formods"), "templates", "DW.yaml") 
#   
# )
# 
# for(yaml in names(yamls)){
# # cat("``` yaml")
#   cat(paste(readLines(file.path(system.file(package="formods"), "templates", "formods.yaml"))), collapse="\n")
# }
```
